\input{111pre}
\begin{document}
\MYTITLE{Examination 1 Study Guide \\ Delivered: Tuesday, February 23, 2016 \\ Examination: Thursday, February 25, 2016, 9:30 am}

\vspace*{-.35in}
\section*{Introduction}
\vspace*{-.05in}

This course will have its first examination on Thursday, February 25, 2016 from 9:30 to 10:45 am. The examination will
be ``closed notes'' and ``closed book'' and it will cover the following content. Please review the ``Course Schedule''
on the Web site for the course to see the content and slides that we have covered in the first modules. You may post
questions about this material to Slack.

\vspace*{-.05in}

\begin{itemize}

  \itemsep 0in

  \item Chapters One through Two in DSPP (i.e., the fundamentals of distributed systems).

  \item Chapters Three through Four in DSPP (i.e., the use of processes and remote communication).

  \item Your class notes, class activities, lecture slides, and the first three laboratory assignments.

  \item Knowledge of the basic commands necessary for using the Ubuntu operating system, {\tt git} and Bitbucket; basic
    understanding of how to use \LaTeX~and Markdown for scientific writing.

\end{itemize}

\vspace*{-.05in}

\noindent The examination will include a mix of questions that will require you to draw and/or comment on a diagram,
write a short answer, explain and/or write a source code segment, or give and comment on a list of concepts or points.
The emphasis will be on the following list of illustrative subjects. Please note that this list is not exhaustive ---
rather it is designed to suggest representative topics.

\vspace*{-.05in}
\begin{itemize}

  \itemsep 0in

  \item The state-of-the-art and the key challenges within the field of distributed systems, with a focus on the types
    of transparency and the challenges associated with their provision.

  \item The basic evaluation metrics for distributed systems (e.g., response time and throughput), how they would be
    calculated, and their overall relationship to properties like scalability.

  \item The architectural styles for implementing a distributed system. Specifically, an understanding of how to
    organize clients and servers to ensure, for instance, high reliability and performance.

  \item The trade-offs associated with the use of threads, processes, and light-weight processes in the context of
    clients and servers running in both centralized and distributed computing systems.

  \item The benefits and drawbacks from using resource virtualization in distributed systems.

  \item An understanding of the changing roles of clients and servers (e.g., in the X Window system).

  \item The trade-offs associated with designs for the servers (e.g., stateful versus stateless) and code migration
    techniques (e.g., weak mobility versus strong mobility) in a distributed system.

  \item The protocols commonly used in a distributed system (e.g., HTTP, FTP, and SFTP).

  \item All of the key steps associated with performing remote procedure calls (RPCs).

  \item The similarities and differences between communication with sockets and RPCs or RMIs.

  \item The different types of communication in a distributed system (e.g., multicast and unicast).

  \item Lessons learned from the laboratory assignments in which you implemented and experimentally evaluated
    distributed systems either individually or in a team-based setting.

\end{itemize}

\section*{Examination Policies}

\vspace*{-.05in}
\noindent Minimal partial credit may be awarded for the questions that require a student to write a short answer. You
are strongly encouraged to write short, precise, and correct responses to all of the questions. When you are taking the
examination, you should do so as a ``point maximizer'' who first responds to the questions that you are most likely to
answer correctly for full points. Please keep the time limitation in mind as you are absolutely required to submit the
examination at the end of the class period unless you have written permission for extra time from a member of the
Learning Commons. Students who do not submit their examination on time will have their overall point total reduced.
Please see the course instructor if you have questions about any of these policies.

\vspace*{-.2in}
\section*{Review the Honor Code}
\vspace*{-.1in}

\noindent Students are required to fully adhere to the Honor Code during the completion of this examination. More
details about the Allegheny College Honor Code are provided on the syllabus. Students are strongly encouraged to
carefully review the full statement of the Honor Code before taking \mbox{this test}.

\noindent The following provides you with a review of Honor Code statement from the course syllabus:

The Academic Honor Program that governs the entire academic program at Allegheny College is described in the Allegheny
Academic Bulletin.  The Honor Program applies to all work that is submitted for academic credit or to meet non-credit
requirements for graduation at Allegheny College.  This includes all work assigned for this class (e.g., examinations,
laboratory assignments, and the final project).  All students who have enrolled in the College will work under the Honor
Program.  Each student who has matriculated at the College has acknowledged the following pledge:

\vspace*{-.11in}
\begin{quote}
  I hereby recognize and pledge to fulfill my responsibilities, as defined in the Honor Code, and to maintain the
  integrity of both myself and the College community as a whole.
\end{quote}
\vspace*{-.11in}

\vspace*{-.2in}
\section*{Strategies for Studying}
\vspace*{-.1in}

As you study for this examination, you are encouraged to form study groups with individuals who were previously, during
a laboratory session, a member of one of your software development and empirical study teams. You can collaborate with
these individuals to ensure that you understand all of the key concepts mentioned on this study guide. Additionally,
students are encouraged to create a Slack channel that can host questions and answers that arise as you continue to
study for the test.  Even though the course instructor will try to, whenever possible, answer review questions that
students post in this channel, you are strongly encouraged to answer the questions posted by your colleagues as this
will also help you to ensure that you fully understand the material.

When studying for the test, don't forget that the Web site for our course contains mobile-ready slides that will provide
you with an overview of the key concepts that we discussed in the first modules. You can use the color scheme in the
slides to notice points where we, for instance, completed an in-class activity, discussed a key point, or made reference
to additional details available in the DSPP textbook. Finally, students are strongly encouraged to schedule a meeting
during the course instructor's office hours so that we can resolve any of your questions about the material and ensure
that you have the knowledge and skills necessary for doing well on this examination. Remember, while the test is taken
individually, your review for it can be done collaboratively!

\end{document}
