%!TEX root=cs440S2014-syllabus.tex
% mainfile: cs440S2014-syllabus.tex 
\input{syllabuspre.tex}

\usepackage[compact]{titlesec}

\begin{document}
\MYTITLE{Syllabus}
\MYHEADERS{Syllabus}{}

\subsection*{Course Instructor}
Dr.\ Gregory M.\ Kapfhammer\\
\noindent Office Location: Alden Hall 108 \\
\noindent Office Phone: +1 814-332-2880 \\
\noindent Email: \url{gkapfham@allegheny.edu} \\
\noindent Twitter: \url{@GregKapfhammer} \\
\noindent Web Site: \url{http://www.cs.allegheny.edu/sites/gkapfham/} 

\subsection*{Instructor's Office Hours}

\begin{itemize}
	\itemsep 0em
	\item Monday: 4:30 pm -- 5:30 pm (15 minute time slots)
	\item Wednesday: 1:00 pm -- 4:00 pm (15 minute time slots)
	\item Thursday: 1:00 pm -- 3:00 pm (15 minute time slots)
	\item Friday: 9:00 -- 10:00 am (10 minute time slots) {\em and} \\ \hspace*{.49in} 1:00 pm -- 2:00 pm (5 minute time slots)
\end{itemize}

\noindent 
To schedule a meeting with me during my office hours, please visit my Web site and click the ``Schedule'' link
in the top right-hand corner. Now, you can browse my office hours or schedule an appointment by clicking the correct
link and then reserving an open time slot. From 3:30 -- 4:30 pm on Tuesdays, I also will be available for walk-in
technical consultations in Alden Hall 101.  

\subsection*{Course Meeting Schedule}

Lecture, Discussion, and Group Work: Tuesday, Thursday, 11:00 am -- 12:15 pm\\
Laboratory Session: Monday, 2:30 pm -- 4:20 pm \\
Final Examination: Thursday, May 1, 2014 at 7:00 pm

\subsection*{Course Catalogue Description}

\begin{quote}

A study of operating systems and their design and implementation. Topics include the evolution of modern computer
operating systems, basic components of an operating system, and hardware issues of importance to operating system design
and implementation. Consideration is given to processes and threads, resource (processor, memory, device, and software)
management, process and device scheduling, deadlocks, input/output, file systems, and system security. A representative
operating system is examined in detail. One laboratory per week. Prerequisite: Computer Science 210. 

\end{quote}

\subsection*{Course Objectives}

The design, implementation, and use of operating systems involves the application of many interesting
theories, techniques, methodologies, and tools.  This course has the objective to: 

\begin{enumerate}

  \itemsep 0in

  \item Provide an overview of the nature and functions of operating systems.

  \item Study the relationship between computer hardware and operating systems.
  
  \item Enhance the understanding of the services that operating systems provide to users.

  \item Explore fundamental operating system concepts (e.g., processes, address spaces, and files). 

  \item Study the algorithms in operating systems (e.g., memory management and scheduling).

  \item Examine, in detail, the design of the important modules in an operating system.

  \item Develop a ``big picture'' understanding of the overall design of an operating system.

  \item Provide an introduction to multi-processor systems and distributed systems.

  \item Develop a basic understanding of security and protection issues in operating systems.

  \item Enhance knowledge of the Linux operating system and its associated tools.

\end{enumerate}

\vspace*{-.075in} 
\noindent 
Throughout the semester students also will enhance their ability to write and present ideas
about operating systems in a clear and compelling fashion.  Students will gain practical experience in
the design, implementation, and analysis of operating systems during laboratory sessions and a final project. Finally,
students will develop a richer understanding of the fascinating connections between operating systems and other
disciplines in the social and natural sciences and the humanities.

% \subsection*{Performance Objectives}

\subsection*{Required Textbook}

\noindent{\em Modern Operating Systems}. Andrew S.\ Tanenbaum.
Third Edition, ISBN-10: 0136006639, ISBN-13: 978-0136006633, 1104 pages, 2007. \\
(References to the textbook are abbreviated as ``MOS'' on the course Web site).

\noindent
Students who want to improve their technical writing skills may consult the following books.

\noindent{\em BUGS in Writing: A Guide to Debugging Your Prose}. Lyn Dupr\'e. Second Edition,  ISBN-10: 020137921X,
ISBN-13: 978-0201379211, 704 pages, 1998.

\noindent{\em Writing for Computer Science}.  Justin Zobel. Second Edition,  ISBN-10: 1852338024, ISBN-13:
978-1852338022, 270 pages, 2004.

\noindent
Along with reading the required textbook, you will be asked to study additional articles from a wide variety of
conference proceedings, journals, and the popular press.

\subsection*{Class Policies}

\subsubsection*{Grading}

The grade that a student receives in this class will be based on the following categories. All percentages are
approximate and, if the need to do so presents itself, it is possible for the assigned percentages to change during the
academic semester. 

\begin{center}
  \begin{tabular}{ll}
    Class Participation and Instructor Meetings & 5\% \\
    First Examination & 15\% \\
    Second Examination & 15\% \\
    Final Examination & 20\% \\
    Laboratory Assignments & 30\% \\
    Final Project & 15\%
  \end{tabular}
\end{center}

% \vspace*{-.1in}
\newpage
\noindent
These grading categories have the following definitions:
\vspace*{-.1in}

\begin{itemize}

	\item {\em Class Participation and Instructor Meetings}: All students are required to actively participate
	  during all of the class sessions. Your participation will take forms such as answering questions about the
	  required reading assignments, asking constructive questions of your group members, giving presentations, and
	  leading a discussion session. Furthermore, all students are required to meet with the course instructor during
	  office hours for a total of thirty minutes throughout the Spring 2014 semester.  These meetings must be scheduled
	  through the course instructor's reservation system and documented on a meeting record that you submit on the
	  day of the final examination. A student will receive an interim and final grade for this category.

	\item {\em First and Second Examinations}: The first and second interim examinations will cover all of the material
	  in their associated module(s).  While the second examination is not cumulative, it will assume that a student has a
	  basic understanding of the material that was the focus of the first examination.  The date for the first and
	  second examinations will be announced at least one week in advance of the scheduled date.  Unless prior
	  arrangements are made with the course instructor, all students will be expected to take these examinations on the
	  scheduled date and complete the tests in the stated period of time.

	\item {\em Final Examination}: The final examination is a three-hour cumulative test.  By enrolling in this
	  course, students agree that, unless there are extenuating circumstances, they will take the final examination
	  at the date and time stated on the first page of the syllabus.

	\item {\em Laboratory Assignments}: These assignments invite students to explore different techniques for
	  designing, implementing, evaluating, and documenting software solutions to challenging problems in the field
	  of operating systems.  Many of the assignments will require students to write programs, conduct experiments,
	  and collect, analyze, and write about data sets.  To best ensure that students are ready to develop software
	  in both other classes at Allegheny College and after graduation, students will complete assignments either on
	  an individual basis or in teams.  When teamwork is required, the instructor will assign individuals to teams.
	
	\item {\em Final Project}: This project will present you with the description of an operating systems problem
	  and ask you to design and implement a correct and carefully evaluated solution. Completion of the final
	  project will require you to apply all of the knowledge and skills that you have accumulated during the course
	  of the semester to solve a problem and, whenever possible, make your solution and results publicly available
	  in a free and open fashion.
		
\end{itemize}

\subsubsection*{Assignment Submission}

All assignments will have a stated due date. The printed version of the assignment is to be turned in at the beginning
of the class on that due date; the printed materials must be dated and signed with the Honor Code pledge of the
student(s) completing the work.  Late assignments will be accepted for up to one week past the assigned due date with a
15\% penalty. All late assignments must be submitted at the beginning of the session that is scheduled one week after
the due date. Unless special arrangements are made with the course instructor, no assignments will be accepted after the
late deadline. For any assignment completed in a group, students also must turn in a one-page document that describes
each group member's contribution to the submitted deliverables.  

\subsubsection*{Attendance}

It is mandatory for all students to attend all of the class and laboratory sessions. If you will not be able to attend a
session, then please see the course instructor at least one week in advance to describe your situation.  Students who
miss more than five unexcused classes, laboratory sessions, or group project meetings will have their final grade in the
course reduced by one letter grade. Students who miss more than ten of the aforementioned events will automatically fail
the course.

% \subsection*{Laboratory Attendance Policy}
% 
% In order to acquired the proper skills in technical writing, critical reading, and the presentation of technical
% material, it is essential for students to have hands-on experience in a laboratory. Therefore, it is mandatory for all
% students to attend the laboratory sessions. If you will not be able to attend a laboratory, then please see the course
% instructors at least one week in advance in order to explain your situation. Students who miss more than two unexcused
% laboratories will have their final grade in the course reduced by one letter grade.  Students who miss more than four
% unexcused laboratories will automatically fail the course.
% 

\subsubsection*{Use of Laboratory Facilities}

Throughout the semester, we will experiment with many different software tools that computer scientists use during the
design, implementation, evaluation, and interaction with operating systems.  The course instructor and the
department's systems administrator have invested a considerable amount of time to ensure that our laboratories support
the completion of both the laboratory assignments and the final project.  To this end, students are required to complete
all assignments and the final project while using the department's laboratory facilities. The course instructor and the
systems administrator normally do not assist students in configuring their personal computers.

\subsubsection*{Class Preparation}

% The study of the computer science discipline is very challenging.  Students in this class will be challenged to learn
% the principles and practice of software development.  During the coming semester even the most diligent student will
% experience times of frustration when they are attempting to understand a challenging concept or complete a difficult
% laboratory assignment.  In many situations some of the material that we examine will initially be confusing : do not
% despair!  Press on and persevere!
% 

In order to minimize confusion and maximize learning, students must invest time to prepare for the class discussions and
lectures.  During the class periods, the course instructor will often pose demanding questions that could require group
discussion, the creation of a program or data set, a vote on a thought-provoking issue, or a group presentation.  Only
students who have prepared for class by reading the assigned material and reviewing the current assignments will be able
to effectively participate in these discussions.  More importantly, only prepared students will be able to acquire the
knowledge and skills that are needed to be successful in both this course and the field of computer science.  In order
to help students remain organized and effectively prepare for classes, the course instructor will maintain a class
schedule with reading assignments and presentation slides.   During the class sessions students will also be required to
download, use, and modify programs and data sets that are made available through the course Web site.  

\subsubsection*{Email}

Using your Allegheny College email address, I will sometimes send out class announcements about matters such as
assignment clarifications or changes in the schedule. It is your responsibility to check your email at least once a day
and to ensure that you can reliably send and receive emails. This class policy is based on the following statement in
{\em The Compass}, the College's student handbook.

% \vspace*{-.1in}
% \newpage
\begin{quote}
``The use of email is a primary method of communication on campus. \ldots
All students are provided with a campus email account and address while
enrolled at Allegheny and are expected to check the account on a regular
basis.'' 
\end{quote}
\vspace*{-.15in}

\subsubsection*{Disability Services}

The Americans with Disabilities Act (ADA) is a federal anti-discrimination statute that provides comprehensive civil
rights protection for persons with disabilities.  Among other things, this legislation requires all students with
disabilities be guaranteed a learning environment that provides for reasonable accommodation of their disabilities.
Students with disabilities who believe they may need accommodations in this class are encouraged to contact Disability
Services at 332-2898.  Disability Services is part of the Learning Commons and is located in Pelletier Library.
Please do this as soon as possible to ensure that approved accommodations are implemented in a timely fashion.

\subsubsection*{Honor Code}

The Academic Honor Program that governs the entire academic program at Allegheny College is described in the Allegheny
Course Catalogue.  The Honor Program applies to all work that is submitted for academic credit or to meet non-credit
requirements for graduation at Allegheny College.  This includes all work assigned for this class (e.g., examinations,
laboratory assignments, and the final project).  All students who have enrolled in the College will work under the Honor
Program.  Each student who has matriculated at the College has acknowledged the following pledge:

\vspace*{-.1in}
\begin{quote}
I hereby recognize and pledge to fulfill my responsibilities, as defined in the Honor Code, and to maintain the
integrity of both myself and the College community as a whole.  
\end{quote}
\vspace*{-.15in}

\noindent It is recognized that an important part of the learning process in any course, and particularly one in
computer science, derives from thoughtful discussions with teachers and fellow students.  Such dialogue is encouraged.
However, it is necessary to distinguish carefully between the student who discusses the principles underlying a problem
with others and the student who produces assignments that are identical to, or merely variations on, someone else's
work.  While it is acceptable for students in this class to discuss their programs, data sets, and reports with their
classmates, deliverables that are nearly identical to the work of others will be taken as evidence of violating the
\mbox{Honor Code}.

\subsection*{Welcome to an Adventure in Operating Systems}

In reference to software, Frederick P.\ Brooks, Jr.\ wrote in chapter one of {\em The Mythical Man Month}, ``The magic
of myth and legend has come true in our time.'' Software is a pervasive aspect of our society that changes how we think
and act. Moreover, the operating system is one of the crucial software components that enables people to interact with a
wide range of servers, desktop workstations, and mobile and embedded devices.  As such, efficient and correct operating
systems have the potential to positively influence the lives many of people.  Moreover, the design, implementation,
evaluation, documentation, and use of operating are exciting and rewarding activities!  At the start of this class, I
invite you to pursue this adventure in operating systems with enthusiasm and vigor.

\end{document}
